Structure Learning for $\VAR(p)$ Models

### 2019-04-30 paper.tex
Expanding j ->PW i necessary conditions proposition.

  \begin{equation}
    \label{eqn:ancestor_expansion}
    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\A_{ik}(z)v_k(t),
  \end{equation}

@@

  \begin{align*}
    x_i(t) &= (I - \B(z))^{-1}\\
    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
    &= \A_{ii}(z) v_i(t) + \sum_{j \in \anc{i} \setminus \{i\}} \A_{ij}(z) v_j(t)\\
  \end{align*}
\end{proof}

@@

  \begin{enumerate}
    \item{$j \in \anc{i}$}
    \item{$\anc{i} \cap \anc{j} \ne \emptyset$}
  \end{enumerate}
\end{proposition}
\begin{proof}
  We will show the contrapositive.  Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have

  \begin{align*}
    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
    &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
    &+ \sum_{k \in \anc{i} \setminus i}\sum_{\ell \in \anc{j} \setminus j}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
  \end{align*}

@@

  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.
\end{proof}

We can state some interesting corollaries of this result.

@@

@@

  but this is impossible since with probability one $v_i(t) \not \in \H_{t - 1}^{(i)}$.
\end{proof}

>>>>

  \begin{equation}
    \label{eqn:ancestor_expansion}
    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t),
  \end{equation}

@@

  \begin{align*}
    x_i(t) &= (I - \B(z))^{-1}\\
    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)\\
  \end{align*}
\end{proof}

@@

  \begin{enumerate}
    \item{$j \in \anc{i}$}
    \item{$\anc{i} \cap \anc{j} \ne \emptyset$.}
  \end{enumerate}

  \hl{Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.  [Must still prove this.]}
\end{proposition}
\begin{proof}
  We will show the contrapositive.  Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have

  \begin{align*}
    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
    &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
  \end{align*}

@@

  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.

  For the second part of the theorem we exclude possibility $(1)$, namely, $j \not \in \anc{i}$ and suppose by way of contradiction that for every $u \in \anc{i} \cap \anc{j}$ 
\end{proof}

An interesting corollary is the following:

@@

\begin{proposition}
  If $\gcg$ is a strongly causal DAG and $i \pwgc j$ and $j \pwgc i$, then $i \not\in \pa{j}$ and $j \not\in \pa{i}$.  That is, a pairwise bidirectional edge indicates the absense of any edge in $\gcg$.
\end{proposition}
\begin{proof}
  By way of contradiction, suppose that $i \in \pa{j}$ and note that the case $j \in \pa{i}$ is symmetric.  We consider the two possibilities allowed by proposition \ref{prop:ancestor_properties} for $j \pwgc i$:  Firstly $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic; \hl{Secondly} $u \in \anc{i} \cap \anc{j}$ \hl{leads to a contradiction} since there must be multiple paths, namely $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow j$ and $u \rightarrow \cdots \rightarrow j$ in $\gcg$, but this isn't possible since $\gcg$ is strongly causal, therefore we cannot have $i \in \pa{j}$. \hl{This requires a strengthening of proposition 3!}
\end{proof}

@@

  but this is impossible since with probability one $x_i(t) \not \in \H_{t - 1}^{(i)}$.
\end{proof}
